Depth and thickness of tectonic tremor in the northeastern Olympic Peninsula

### 2021-06-03 paper/agujournaltemplate.tex
Annotate cross-sections

\begin{abstract}
Tectonic tremor has been explained as a swarm of low-frequency earthquakes, which are located on a narrow fault at the plate boundary. However, due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source and the distance of the source to the plate interface with great precision. The thickness of the tremor region is not well constrained either. The tremor may be located on a narrow fault like the low-frequency earthquakes, or distributed through a few kilometers wide low shear-wave velocity layer in the upper oceanic crust, which is thought to be a region with high pore-fluid pressure. In this study, we use the cross-correlation of the horizontal and vertical components of seismic data recorded by small-aperture arrays in the Olympic Peninsula, Washington to compute the depth and thickness of the tremor. We find that the source of the tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick. The source of the tremor is thus distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located in the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a layer distributed above and under the plate boundary.
\end{abstract}

@@

The data were collected during the 2009-2010 Array of Arrays experiment. Eight small-aperture arrays were installed in the northeastern part of the Olympic Peninsula, Washington. The aperture of the arrays was about 1 km, and station spacing was a few hundred meters. The arrays were around 5 to 10 km apart from each other (Figure 1). Most of the arrays recorded data  for most of  a year, between June 2009 to September 2010, and captured the main August 2010 ETS event. The arrays also recorded the August 2011 ETS event with a slightly reduced number of stations. ~\citeA{GHO_2012} used a multibeam backprojection (MBBP) technique to detect and locate tremor. They bandpass filtered the vertical component between 5 and 9 Hz and divided the data into one-minute-long time windows. They performed beam forming in the frequency domain at each array to determine the slowness vectors, and backprojected the slownesses through a 3-D wavespeed model ~\cite{PRE_2003} to locate the source of the tremor for each time window. This produced 28902 tremor epicenters for one-minute-long time windows during June, 2009 - September, 2010 and 5600 epicenters during August - September, 2011.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 2.5cm 0cm 9.5cm},clip]{figures/arrays_location.eps}
\caption{Map showing the location of the eight arrays (black triangles) used in this study. Grey dots are the locations of the source of the tremor recorded by the arrays. Inset shows the study area with the box marking the area covered in the main map. Contour lines represent a model of the depth of the plate interface ~\cite{MCC_2006}.}
\label{pngfiguresample}
\end{figure}

@@

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute-long time windows corresponding to all the tremor epicenters located within the grid cell. For each one-minute-long time window, we downloaded the seismic data for each seismic station of the array. Then, for each seismic station and each channel, we detrended the data, tapered the first and last 5 seconds of the data with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross-correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross-correlation functions over all the seismic stations of the array. We experimented with a linear stack, a nth-root stack, and a phase-weighted stack ~\cite{SCH_1997}. We then computed the envelopes of the stacked cross-correlation functions. Figure 2 shows an example of the envelopes of the cross-correlation functions for the Big Skidder array for the 110 one-minute-long time windows when tremor was detected in a 5 km by 5 km grid cell located 7 kilometers southwest of the array. We can see that for about half of the tremor windows, there is a peak in the cross-correlation at about 4.85 s. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we assume that this peak corresponds to the time lag between the arrival of a direct P-wave and a direct S-wave. We then stacked the cross-correlation functions over all the one-minute-long time windows. Again, we experimented with a linear stack, a nth-root stack, and a phase-weighted stack. We assume that the time of the maximum absolute value of the peak of the stack is the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave.

@@

Only about half of the envelopes of the cross-correlation functions have a distinct peak that coincides with the peak in the stacked cross-correlation. The other cross-correlations functions show either a distinct peak at another time lag, or no clearly visible peak. This may be either because the source of the tremor during the corresponding one-minute-long time window was mislocated, or because the signal-to-noise ratio is too low. To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match well the stacked cross-correlation, and the ones that do not match it well. For each one-minute-long time window, we took the envelope of the cross-correlation function between the horizontal and the vertical component, and cross-correlated it with the stack over all the time windows. From this last cross-correlation function, we then computed its maximum absolute value, its value at time 0, and the time lag at which it takes its maximum absolute value. For each one-minute-long time window, we also took the cross-correlation function between the horizontal and the vertical component and computed the ratio between the amplitude of the peak to the root mean square. We did this for both the East-West component and the North-South component. Each one-minute-long time window is thus associated to eight values of quality criteria. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose the number of clusters R, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the cross-correlation functions over all the one-minute-long time windows belonging to the cluster using a phase-weighted stack. We always use two clusters. We try using more clusters, but it did not improve the final stack. Figure 3 shows the stack of the envelopes of the cross-correlation for each of the clusters compared to the original stacked cross-correlation for all the one-minute-long time windows. The clustering has improved the amplitude of the peak for one of the clusters, and made the peak nearly disappear for the other cluster.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_stackcc.eps}
\caption{Stack of the cross-correlation functions over all the 110 time windows from Figure 2. We used a phase-weighted stack for the stack over the stations and for the stack over the time windows. The black line is the stack of the cross-correlation functions over all the time windows. The red line on the left panels is the stack of the envelopes of the cross-correlation functions over the time windows in the first cluster (which contains the time windows that do not fit well with the stack), and the blue line on the right panels is the stack of the envelopes of the cross-correlation functions over the time windows in the second cluster (which contains the time windows that fit well with the stack). Top panels are the cross-correlation of the EW component with the vertical component, and bottom panels are the cross-correlation of the NS component with the vertical component. The stacked cross-correlation function has been cut between 2 and 8 seconds to focus on the time lags around the peak.}
\label{pngfiguresample}
\end{figure}

We did this analysis for grid cells located in a 50 km by 50 km area centered on each of the eight arrays. We thus have 11 * 11 * 8 = 968 values of the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave. We first assumed that the epicenter is at the center of the grid and determined the depth that satisfies the observed S-minus-P time for a 1D velocity model. The velocity models are taken from the 3D velocity model from ~\citeA{MER_2020}. For each array, we looked for the closest grid point (in terms of latitude and longitude) in the 3D model by ~\citeA{MER_2020} and used the corresponding  layered model to compute the depth. We thus used 8 different 1D velocity models corresponding to the 8 arrays. This allows us to take into account the variations of the Poisson's ratio in the East-West direction.

@@

Finally, the data from some of the arrays are very noisy, which makes it hard to see a signal emerging when stacking over the one-minute-long time windows. We chose to keep only the locations for which the ratio between the maximum value of the stack of the envelopes of the cross-correlation functions to the root mean square is higher than 5. We compute the root mean square of the cross-correlation for a time lag between 12 and 14s because we do not expect any reflected wave to arrive that late after a direct wave. The corresponding stack of the envelopes of the cross-correlation functions are shown in Figure 4 for the East-West component and the North-South component. The grey vertical line shows the theoretical time lag between the arrival of the direct P-wave and the arrival of the direct S-wave using the 1D velocity model and the plate boundary model from ~\citeA{MCC_2006}, the dashed line corresponds to the slab model from ~\citeA{PRE_2003}. There is a good agreement between the timing of the peak and the theoretical time lag for most of the locations of the source of the tremor.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={2.5cm 0.5cm 5cm 1cm},clip]{figures/BS_PWS_PWS_0.eps}
\caption{Stack of the envelopes of the cross-correlation signals for different positions of the tremor source relative to the Big Skidder array showing the cross-correlation between the East-West and the vertical components (red lines) and between the North-South and vertical components (blue lines). The theoretical S-minus-P times (grey plain line for the McCrory model, grey dashed line for the Preston model) are generally in good agreement with the peaks in cross-correlation functions. The location of the source of the tremor varies from west to east (left to right) and from south to north (bottom to top). The numbers next to each graph indicate the number of one-minute-long time windows in the best cluster. If number is colored in black, this location is kept in the following maps, if the number is colored in grey, this location is not kept in the maps. For clarity, we plotted only the part of the cross-correlation signal corresponding to timelags between 2 and 8 seconds.}
\label{pngfiguresample}
\end{figure}

@@

%The method used by ~\citeA{GHO_2012} to locate the tremor is based on a multibeam-backprojection (BMMP) technique, which used only vertical channels to compute the slowness vectors, and then a 1D P-wave velocity model for the backprojection. To correct for possible mislocation of the tremor, we relocated the source of the tremor using the tremor catalog from the Pacific Northwest Seismic Network (\url{https://pnsn.org/tremor}) and computed the depth of the tremor using the new relocated latitude and longitude for each of the grid points. \\

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/uncertainty_PWS_PWS.eps}
\caption{Map of the uncertainty on the depth of the source of the tremor based on the width of the stack of the envelopes of the cross-correlation functions. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/depth_PWS_PWS.eps}
\caption{Map of the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The three black lines indicate the positions of the cross-sections shown in Figures 7 to 9. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_1.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_2.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the middle black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_3.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The black line corresponds to the plate boundary profile of ~\citeA{MCC_2006} along the northern most black line shown on Figure 6, and the dashed line to the plate boundary profile of ~\citeA{PRE_2003}. The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the McCrory model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. Tremor located below the plate boundary is blue, and above is red. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the Preston model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}). Tremor located below the plate boundary is blue, and above is red. Locations marked by squares are obtained from seismic data recorded by the Burnt Hill array, stars are associated with Big Skidder, inverted triangles with Danz Ranch, large circles with Gold Creek, diamonds with Port Angeles, and triangles with Three Bumps.}
\label{pngfiguresample}
\end{figure}

@@

Previous studies have shown evidence for seismic anisotropy near the slab interface ~\cite{NIK_2009} and in the overriding continental crust ~\cite{CAS_1996}. If the source of the tremor is located in or under the anisotropic layer, the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave could be different whether we consider the East-West component of the North-South component. To verify whether possible anisotropy could affect our results, we computed the time difference between the timing of the maximum of the stacked of the envelopes of the cross-correlation functions between the East-West component and the vertical component, and the timing of the maximum of the stack of the envelopes of the cross-correlation functions between the North-South component and the vertical component. We also computed the associated difference in depth of the source of the tremor (see Figure 12) for all arrays and all locations of the tremor. The difference in time is generally less than 0.25 s while the difference in depth is generally lower than 2 kilometers, which is less than the uncertainty on the depth of the source of the tremor. As the time difference between the East-West component and the North-South component due to anisotropy is expected to vary depending on the relative position of the array compared to the source of the tremor, we also plotted on Figure 12 the time difference as a function of the distance from source to array, and as a function of the azimuth, as well as the corresponding depth difference. The time difference does not increase with the distance from the source to the array, and there is no seismic path orientation that gives bigger time difference. Thus, seismic anisotropy does not seem to have a significant effect on the time lags between the arrival of the direct P-wave and the arrival of the direct S-wave, and can be neglected. ~\citeA{NIK_2009} observed anisotropy in the low-velocity layer beneath station GNW, located in the eastern Olympic Peninsula, farther south than the eight seismic arrays used in this study. If the source of the tremor is located above the low-velocity layer, we do not expect seismic anisotropy to introduce a significant effect on the time lags measured with the East-West component compared with the time lags measured with the North-South component. ~\citeA{NIK_2009} locate the low-velocity layer above the plate boundary, in the lower continental crust, which is shallower than the location given by ~\citeA{BOS_2013}, who locates the low-velocity layer in the upper oceanic crust. However, even if a low-velocity layer with 5 \% anisotropy is located in the lower continental crust as indicated by ~\citeA{NIK_2009}, the resulting difference in time lags should not be higher than a few tens of seconds, and the resulting difference in depth should not be higher that a few kilometers, which is lower than the uncertainty of the depth of the source of the tremor. ~\citeA{CAS_1996} observed anisotropy in the continental crust, especially in the upper 20 kilometers. However, the anisotropy resulted in time delays of the seismic waves of no more than 0.32s for deep earthquakes (40-60 km depth) and nor more than 0.20s for shallow earthquakes (15-30 km depth). The corresponding difference in depth of the source of the tremor would thus be less than the uncertainty.

>>>>

\begin{abstract}
Tectonic tremor has been explained as a swarm of low-frequency earthquakes, which are located on a narrow fault at the plate boundary. However, due to the lack of clear impulsive phases in the tremor signal, it is difficult to determine the depth of the tremor source and the distance of the source to the plate interface with great precision. The thickness of the tremor region is also not well constrained. The tremor may be located on a narrow fault like the low-frequency earthquakes appear to be, or distributed over a few kilometers wide low shear-wave velocity layer in the upper oceanic crust, which is thought to be a region with high pore-fluid pressure. Lag times of peaks in the cross-correlation of the horizontal and vertical components of tremor seismograms, recorded by small-aperture arrays in the Olympic Peninsula, Washington, are interpreted to be S minus P times.  Tremor depths are estimated from these S minus P times using epicenters from a tremor catalog. The tremor is located close to the plate boundary in a region no more than 2-3 kilometers thick and is very close to the depths of low-frequency earthquakes. The tremor is distributed over a wider depth range than the low-frequency earthquakes. However, due to the uncertainty on the depth, it is difficult to conclude whether the source of the tremor is located at the top of the subducting oceanic crust, in the lower continental crust just above the plate boundary, or in a narrow zone at the plate boundary.
\end{abstract}

@@

The data were collected during the 2009-2011 Array of Arrays experiment. Eight small-aperture arrays were installed in the northeastern part of the Olympic Peninsula, Washington. The aperture of the arrays was about 1 km, and station spacing was a few hundred meters. Arrays typically had ten 3-component seismometers, augmented by an additional 10 vertical-only sensors during the 2010 ETS event.  The arrays were around 5 to 10 km apart from each other (Figure 1). Most of the arrays recorded data  for most of  a year, between June 2009 to September 2010, and captured the main August 2010 ETS event. The arrays also recorded the August 2011 ETS event with a slightly reduced number of stations. ~\citeA{GHO_2012} used a multibeam backprojection (MBBP) technique to detect and locate tremor. They bandpass filtered the vertical component between 5 and 9 Hz and divided the data into one-minute-long time windows. They performed beam forming of vertical component data in the frequency domain at each array to determine the slowness vectors, and backprojected the slownesses through a 3-D wavespeed model ~\cite{PRE_2003} to locate the source of the tremor for each time window. This produced 28902 tremor epicenters for one-minute-long time windows during June, 2009 - September, 2010 and 5600 epicenters during August - September, 2011.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 2.5cm 0cm 9.5cm},clip]{figures/arrays_location.eps}
\caption{Map showing the location of the eight arrays (black triangles) used in this study and tremor locations (grey dots) located using these arrays ~\cite{GHO_2012}. Inset shows the study area with the box marking the area covered in the main map. Contour lines represent a model of the depth of the plate interface ~\cite{MCC_2006}.}
\label{pngfiguresample}
\end{figure}

@@

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute-long time windows corresponding to all the tremor epicenters located within the grid cell. For each one-minute-long time window, we downloaded the seismic data for each 3-component seismic station of the array. Then, for each seismic station and each channel, we detrended the data, tapered the first and last 5 seconds of the data with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross correlation functions over all the seismic stations of the array. At each step of this method we experimented with a linear stack, a nth-root stack, and a phase-weighted stack (~\cite{SCH_1997} and found that the phase-weighted stack worked best so all the stacks discussed here used the phase-weighted method. We then computed the envelopes of the stacked cross-correlation functions. Figure 2 shows an example of the envelopes of the cross-correlation functions for the Big Skidder array for the 110 one-minute-long time windows when tremor was detected in a 5 km by 5 km grid cell located 7 kilometers southwest of the array. We can see that for about half of the tremor windows, there is a peak in the cross-correlation at about 4.85 s. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we assume that this peak corresponds to the time lag between the arrival of a direct P-wave and a direct S-wave. We then stacked the cross-correlation functions over all the one-minute-long time windows. We assume that the time of the maximum absolute value of the peak of the stack is the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave.

@@

Only about half of the envelopes of the cross-correlation functions have a distinct peak that coincides with the peak in the stacked cross-correlation. The other cross-correlations functions show either a distinct peak at another time lag, or no clearly visible peak. This may be either because the source of the tremor during the corresponding one-minute-long time window was mislocated, or because the signal-to-noise ratio is too low. To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match well the stacked cross-correlations, and the ones that do not match it well. For each one-minute-long time window, we took the envelope of the stack of cross-correlation functions between the horizontal and the vertical components of each station in an array.  We then cross-correlated it with the stack over all the envelopes. From this last cross-correlation function, we then computed its value at zero lag time, its maximum absolute value, and the time lag at which it takes its maximum absolute value and the ratio of the maximum absolute value to the rms of the correlation function.  We considered lag times from X to Y in each of these measures. We did this for both the East-West component and the North-South component. Each one-minute-long time window is thus associated with eight values of quality criteria. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose the number of clusters R, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the cross-correlation functions over all the one-minute-long time windows belonging to the cluster using a phase-weighted stack. We always use two clusters. We try using more clusters, but it did not improve the final stack. Figure 3 shows the stack of the envelopes, as described above, for each of the two clusters as well as a stack over all windows of the stack of cross correlation functions. The clustering has greatly improved the amplitude of the peak for one of the clusters, and made the peak nearly disappear for the other cluster.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0cm 0cm 0cm 0cm},clip]{figures/BS_-05_-05_PWS_PWS_cluster_stackcc.eps}
\caption{Stack of the cross-correlation functions over all the 110 time windows from Figure 2. We used a phase-weighted stack for the stack over the stations and for the stack over the time windows. The black line is the stack over all time windows of the stack of cross-correlation functions over all stations within the array. The red line on the left panels is the stack over all the time windows in the first cluster of the envelopes of the stack of cross-correlation functions over stations within the array.  This corresponds to time windows that do not fit well with the stack. The blue lines on the right panels are the same as the red lines, except it uses time windows in the second cluster (which contains the time windows that fit well with the stack). Top panels are the cross-correlation of the EW component with the vertical component, and bottom panels are the cross-correlation of the NS component with the vertical component. The stacked cross-correlation function has been cut between 2 and 8 seconds to focus on the time lags around the peak.}
\label{pngfiguresample}
\end{figure}

We did this analysis for grid cells located in a 50 km by 50 km area centered on each of the eight arrays. We thus consider 11 * 11 * 8 = 968 values of the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave. We first assumed that the epicenter is at the center of the grid and determined the depth that satisfies that epicenter and the observed S-minus-P. The velocity models are taken from the 3D velocity model from ~\citeA{MER_2020}. For each array, we looked for the closest grid point (in terms of latitude and longitude) in the 3D model by ~\citeA{MER_2020} and used the corresponding layered model of compressional and shear wave speeds to compute the depth. We thus used 8 different 1D velocity models corresponding to the 8 arrays. This allows us to take into account the substantial variations of the Poisson's ratio in the East-West direction.

@@

Finally, the data from some of the arrays are very noisy, which makes it hard to see a signal emerging when stacking over the one-minute-long time windows. We chose to keep only the locations for which the ratio between the maximum value of the stack of the envelopes of the cross-correlation functions to the root mean square is higher than 5. We compute the root mean square of the cross-correlation for a time lag between 12 and 14s because we do not expect any reflected wave to arrive that late after a direct wave. The corresponding stack of the envelopes of the cross-correlation functions are shown in Figure 4 for the East-West component and the North-South component. The grey vertical line shows the theoretical time lag between the arrival of the direct P-wave and the arrival of the direct S-wave using the corresponding 1D velocity model for this array and the plate boundary model from ~\citeA{MCC_2006}, the dashed line corresponds to the slab model from ~\citeA{PRE_2003}. There is a good agreement between the timing of the peak and the theoretical time lag for most of the locations of the source of the tremor.

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={2.5cm 0.5cm 5cm 1cm},clip]{figures/BS_PWS_PWS_0.eps}
\caption{Stack of the envelopes of the cross-correlation signals for different positions of the tremor source relative to the Big Skidder array showing the cross-correlation between the East-West and the vertical components (red lines) and between the North-South and vertical components (blue lines). The theoretical S-minus-P times (grey plain line for the McCrory model, grey dashed line for the Preston model) are generally in good agreement with the peaks in cross-correlation functions. The location of the tremor varies from west to east (left to right) and from south to north (bottom to top). The numbers next to each graph indicate the number of one-minute-long time windows in the best cluster. For clarity, we plotted only the part of the cross-correlation signal corresponding to timelags between 2 and 8 seconds.}
\label{pngfiguresample}
\end{figure}

@@

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/uncertainty_PWS_PWS.eps}
\caption{Map of the uncertainty on the depth of the tremor based on the width of the stack of the envelopes of the cross-correlation functions for arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/depth_PWS_PWS.eps}
\caption{Map of the depth of the source of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. The three black lines indicate the positions of the cross-sections shown in Figures 7 to 9. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles)}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_1.eps}
\caption{Cross-section showing depth of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR} for cross section A-A' shown in Figure 6. The black line corresponds to the plate boundary profile of McCrory (2006) along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of Preston (2003). The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_2.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families for cross section B-B' shown in Figure 6. Symbols are described in caption for Figure 7.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_3.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families for cross section C-C' shown in Figure 6. Symbols are describe in caption for Figure 7.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the McCrory model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. Tremor located below the plate boundary is blue, and above is red. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the Preston model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}). Tremor located below the plate boundary is blue, and above is red. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

Previous studies have shown evidence for seismic anisotropy near the slab interface ~\cite{NIK_2009} and in the overriding continental crust ~\cite{CAS_1996}. If the source of the tremor is located in or under the anisotropic layer, the time lag between the arrival of the direct P-wave and the arrival of the direct S-wave could be different different depending on whether we consider the East-West component or the North-South component. To verify whether possible anisotropy could affect our results, we computed the time difference between the timing of the maximum of the stack of the envelopes of the stacked cross-correlation functions between the East-West component and the vertical component, and between the North-South component and the vertical component. We also computed the associated difference in depth of the source of the tremor (see Figure 12) for all arrays and all locations of the tremor. The difference in time is generally less than 0.25 s while the difference in depth is generally less than 2 kilometers, which is similar to the uncertainty on the tremor depth. As the time difference between the East-West component and the North-South component due to anisotropy is expected to vary depending on the relative position of the array compared to the source of the tremor, we also plotted the time difference as a function of the distance from source to array, and as a function of the azimuth, as well as the corresponding depth difference (Figure 12). The time difference does not increase with the distance from the source to the array, and there is no seismic path orientation that gives bigger time difference. Thus, seismic anisotropy does not seem to have a significant effect on the time lags between the arrival of the direct P-wave and the arrival of the direct S-wave, and can be neglected. ~\citeA{NIK_2009} observed anisotropy in the low-velocity layer beneath station GNW, located in the eastern Olympic Peninsula, south of the eight seismic arrays used in this study. If the tremor is located above the low-velocity layer, we do not expect this source of seismic anisotropy to introduce a significant effect on our measured time lags. ~\citeA{NIK_2009} locate the low-velocity layer above the plate boundary, in the lower continental crust, which is shallower than the location given by ~\citeA{BOS_2013}, who locates the low-velocity layer in the upper oceanic crust. However, even if a low-velocity layer with 5 \% anisotropy is located in the lower continental crust as indicated by ~\citeA{NIK_2009}, the resulting difference in time lags should not be more than a few tenths of a second, and the resulting difference in depth should not be more than 2 kilometers. ~\citeA{CAS_1996} observed anisotropy in the continental crust, especially in the upper 20 kilometers to the north of our arrays. However, the anisotropy resulted in time delays of the seismic waves of no more than 0.32s for deep earthquakes (40-60 km depth) and not more than 0.20s for shallow earthquakes (15-30 km depth). The corresponding difference in tremor depth is similar to the uncertainty.
